The Power of Unbiased Recursive Partitioning:\\ A Unifying View of CTree, MOB, and GUIDE

### 2019-02-27 pkg/partykit/inst/guideliketest/DAGStat2019/slides.Rnw
some touch-ups
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2478 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: d0468c79a26da1f358c42b0dcb7ac6b194b835d5

\documentclass[11pt,t,usepdftitle=false,aspectratio=43]{beamer}
\usetheme[nototalframenumber, noslidenumber]{uibk}

\title{The Power of Unbiased Recursive Partitioning:\\
A Unifying View of CTree, MOB, and GUIDE}
\author{Lisa Schlosser, Torsten Hothorn, Achim Zeileis}
\setbeamerfont{url}{size*={11.5pt}{13pt},series=\mdseries}
%\renewcommand{\mysize}{\fontsize{7.7pt}{9pt}\selectfont}
\URL{https://R-Forge.R-project.org/projects/partykit/} 
%\\
%https://partykit.org}
\headerimage{3}

\setbeamertemplate{caption}{\insertcaption} 
%% includes a replacement for \usepackage{Sweave}
\usepackage{Sweave}
\usepackage{changepage}
\usepackage{amsmath,tikz}
\usetikzlibrary{positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap,trees,tikzmark,decorations.pathreplacing}
%\graphicspath{{plots/}}

@@

\begin{frame}%[fragile]
\frametitle{Simulation Setting}

@@

\begin{frame}%[fragile]
\frametitle{Simulation Setting (Scenario 1)}
True tree structure:

@@

\begin{frame}%[fragile]
\frametitle{Results}
Residuals vs.\ full model scores:

@@

\begin{frame}%[fragile]
\frametitle{Results}
Full factorial analysis (with both parameters varying):
\begin{figure}%[t]
\setkeys{Gin}{width=1\linewidth}
<<3way, fig=TRUE, echo=FALSE, height=4, width=8>>=
xyplot(pval_z1 ~ cat | xi, 
       groups = interaction(bin, res_scores), 
       data = d3way, 
       ylim = c(0,0.5),
       type = c("a"), 
       col = c("red", "blue", "red", "blue"),
       lty = c(1,1,2,2),
       xlab = "Categorization",
       ylab = expression("p-value of"~Z[1]),
       #index.cond = list(c(1,2,3,4)),
       strip = strip.custom(factor.levels = c(expression(xi==0~"and"~delta==0.3), 
                                              expression(xi==0.8~"and"~delta==1))),
       key = list(text = list(levels(interaction(d3way$bin, d3way$res_scores))),
                  lines = list(lty = c(1,1,2,2), 
                               type = "l",
                               col = c("red", "blue", "red", "blue"))))
@
\end{figure}
\end{frame}

\begin{frame}%[fragile]
\frametitle{Simulation Setting (Scenario 2)}
True tree structure:

@@

\begin{frame}%[fragile]
\frametitle{Results}

Residuals vs.\ full model scores:

@@

\begin{frame}%[fragile]
\frametitle{Results}
Pre-pruning vs.\ post-pruning:

@@

\begin{frame}%[fragile]
\frametitle{Recommendations}
Based on the simulation results the following general conclusions can be drawn:
\begin{itemize}
\item Including full model scores is advantageous over using only residuals.
\item Binarizing scores/residuals does not lead to any improvements in performance.
\item Categorizing split variables can facilitate calculations, however, important
effects might be missed unless a binning point is close to the true split point.
\item If the applied statistical test performs well as significance test pre-pruning
is preferable. Otherwise post-pruning can compensate this deficiency at least partly.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Software}
\vspace{0.4cm}
\textbf{Package:} \emph{partykit} available on CRAN\\

@@

\textbf{Reproduction code:} available on R-Forge at\\

>>>>

\documentclass[11pt,t,usepdftitle=false,aspectratio=169]{beamer}
\usetheme[license, url]{uibk}

\title{The Power of Unbiased Recursive Partitioning:\\
A Unifying View of CTree, MOB, and GUIDE}
\author{Lisa Schlosser, Torsten Hothorn, Achim Zeileis}
\setbeamerfont{url}{size*={11.5pt}{13pt},series=\mdseries}
%\renewcommand{\mysize}{\fontsize{7.7pt}{9pt}\selectfont}
\URL{http://www.partykit.org/} 
\headerimage{3}

\setbeamertemplate{caption}{\insertcaption} 
%% includes a replacement for \usepackage{Sweave}
\usepackage{changepage}
\usepackage{amsmath,tikz}
\usetikzlibrary{positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap,trees,tikzmark,decorations.pathreplacing}
%\graphicspath{{plots/}}

@@

\begin{frame}%[fragile]
\frametitle{Simulation}

@@

\begin{frame}%[fragile]
\frametitle{Simulation 1: True tree structure}

@@

\begin{frame}%[fragile]
\frametitle{Simulation 1: Residuals vs.\ full model scores}

@@

\begin{frame}%[fragile]
\frametitle{Simulation 1: Full factorial analysis (vary both)}

\begin{figure}%[t]
\setkeys{Gin}{width=1\linewidth}
<<3way, fig=TRUE, echo=FALSE, height=4, width=8>>=
xyplot(pval_z1 ~ cat | xi, 
       groups = interaction(bin, res_scores), 
       data = d3way, 
       ylim = c(0,0.5),
       type = c("a"), 
       col = c("red", "blue", "red", "blue"),
       lty = c(1,1,2,2),
       xlab = "Categorization",
       ylab = expression("p-value of"~Z[1]),
       #index.cond = list(c(1,2,3,4)),
       strip = strip.custom(factor.levels = c(expression(xi==0~"and"~delta==0.3), 
                                              expression(xi==0.8~"and"~delta==1))),
       key = list(text = list(levels(interaction(d3way$bin, d3way$res_scores))),
                  lines = list(lty = c(1,1,2,2), 
                               type = "l",
                               col = c("red", "blue", "red", "blue"))))
@
\end{figure}
\end{frame}

\begin{frame}%[fragile]
\frametitle{Simulation 2: True tree structure}

@@

\begin{frame}%[fragile]
\frametitle{Simulation 2: Residuals vs.\ full model scores}

@@

\begin{frame}%[fragile]
\frametitle{Simulation 2: Pre-pruning vs.\ post-pruning}

@@

\begin{frame}%[fragile]
\frametitle{Recommendations}

Based on the simulation results the following general conclusions can be drawn:
\begin{itemize}
\item Including full model scores is advantageous over using only residuals.
\item Binarizing scores/residuals does not lead to any improvements in performance.
\item Categorizing split variables can facilitate calculations, however, important
effects might be missed unless a binning point is close to the true split point.
\item If the applied statistical test performs well as significance test pre-pruning
is preferable. Otherwise post-pruning can compensate this deficiency at least partly.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Software}
\vspace{0.4cm}
\textbf{Package:} Available on CRAN at\\

\url{https://CRAN.R-project.org/package=partykit}\\

@@

\textbf{Replication code:} Available on R-Forge at\\
